---
title: "Chi-Squared Tests"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(dplyr)
library(rcompanion)
gradethis_setup()
knitr::opts_chunk$set(echo = FALSE)

```

```{r prep-data}
titanic <- as.data.frame(Titanic) %>% 
  dplyr::filter( Age == "Adult") %>% 
  droplevels() %>% 
  select(Class, Sex, Freq)

tab_titanic <- xtabs(Freq ~ Class + Sex, data = titanic)

```

When we get to a situation with non-continuous $x$ (explanatory) and $y$ (response) variables we move to the $\chi^2$ test instead of doing linear models. 

Here we shall examine a data frame that reports survival of passengers on the ill-fated ship Titanic, it is called `titanic` 

View the data frame.

```{r, chi-0, exercise=TRUE, exercise.setup="prep-data"}

```
```{r, chi-0-solution}
titanic
```
```{r, chi-0-check}
grade_code()
```

```{r, chi-1}
question(
  "Which variable holds the count of survivors?",
  answer("Sex"),
  answer("Class"),
  answer("Freq", correct=TRUE)
)
```

To use the data frame in a $\chi^2$-test we must first tabulate it, we can use `xtabs` and an R formula to do that. The response variable would be our count of survivors. Make sure that the variable of interest is in the rows of the resulting table (Hint: change the order of the explanatory variables in the formula and see what happens).

```{r, chi-2, exercise=TRUE, exercise.setup="prep-data"}
tab_titanic <- xtabs(... ~ ... + ..., data = ...)
tab_titanic
```
```{r, chi-2-solution}
tab_titanic <- xtabs(Freq ~ Sex + Class, data = titanic)
tab_titanic
```
```{r, chi-2-check}
grade_code()
```

Now we have the table in the right format to do a $\chi^2$ test. The `rcompanion` package contains a function called `pairwiseNominalIndependence()`. Use that to assess the whether the difference in number of survivors of different classes is likely the same as others classes.

```{r, chi-3, exercise=TRUE, exercise.setup="prep-data"}
library(...)
pairwiseNominalIndependence(..., method="fdr")
```
```{r, chi-3-solution}
library(rcompanion)
pairwiseNominalIndependence(tab_titanic, method="fdr")
```

```{r, chi-3-check}
grade_code()
```


