---
title: "t-tests"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(palmerpenguins)
library(dplyr)
library(magrittr)
gradethis_setup()
knitr::opts_chunk$set(echo = FALSE)

penguins_small <- select(penguins, species, bill_length_mm) %>% 
  filter(species %in% c('Adelie','Chinstrap')) %>% 
  droplevels()
```


We've seen how ths slope and it's gradient and the statistics generated in a linear model can give us a test for differences that is the same as a $t$-test. Let's use the `penguins` data set to examine some differences in bill size in some penguins.

## Examining the data

Examine the provided data `penguins_small`, use `str()`

```{r, penguins-1, exercise=TRUE}

```

```{r, penguins-1-solution}
str(penguins_small)
```

```{r, penguins-1-check}
grade_code()
```

```{r, penguins-2}
question("How many experimental groups are there?",
         answer("1", message="Two-sample $t$-tests need two groups."),
         answer("2", correct=TRUE,message="Two-sample $t$-tests need two groups."),
         answer("3", message="Two-sample $t$-tests need two groups.")
)
```

Get a summary of the data per group by modifying the code below
```{r, penguins-3, exercise=TRUE}
penguins_small %>% 
  group_by() %>% 
  summarize(mean_bill_length = mean(bill_length_mm),
            sd_bill_length = sd(bill_length_mm)
            )
```
```{r, penguins-3-solution}
penguins_small %>% 
  group_by(species) %>% 
  summarize(mean_bill_length = mean(bill_length_mm, na.rm = TRUE),
            sd_bill_length = sd(bill_length_mm, na.rm = TRUE)
  )
```
```{r, penguins-3-check}
grade_code(glue_incorrect="Remember to group by the right column, and to remove NA values (na.rm=TRUE) in summary functions.")

```

## $t$-tests

Build a linear model suitable for the `penguins_small` data set. Generate a summary.

```{r, penguins-4, exercise=TRUE}
model <- lm()
```
```{r, penguins-4-solution}
model <- lm(bill_length_mm ~ species, data = penguins_small)
summary(model)
```
```{r, penguins-4-check}
grade_code()
```

Recall that the $F$ statistic indicates the strength of relationship between the $x$ and $y$, in other words the size of and fit to the slope, which together indicates a real difference between the groups.

```{r, penguins-5}
question(
  "What is the $F$ statistic between these groups?",
  answer("566.5", correct=TRUE),
  answer("38.7914"),
  answer("10.0424")
)
```

```{r, penguins-6}
question(
  "Is this value a strong $F$ value? Which of these statements is most true?",
  answer("It is low, there isn't evidence of differences.",message="Remember $F$ is a ratio, moving above 1 indicates a relationship. this $F$ value is quite a long way over 1."),
  answer("It is moderate, there is some relationship and some difference between the groups.", message="Remember $F$ is a ratio, moving above 1 indicates a relationship. this $F$ value is quite a long way over 1."),
  answer("Yes, there is a good relationship and likely a real difference", correct=TRUE)
)
```

Remember that the coefficient tells us about the change in $y$ per unit increase in $x$. In our two group situation we set things up so that the distance on the $x$-axis between groups is 1. 

```{r, penguins-8}
question(
  "What is the size difference between the bills of the two penguin types?",
  answer("Chinstrap is bigger than Adelie by 10.0424 cm", message="Check your units"),
  answer("Adelie is bigger than Chinstrap by 38.7914 mm"),
  answer("Adelie is smaller than Chinstrap by 10.0424 mm", correct=TRUE),
 answer("Adelie is smaller than Chinstrap by 38.7914 mm")
  )
```


```{r, penguins-9}
question(
  "What is the $p$-value for the likelihood that the difference of the size observed occurs at chance in a Null Model?",
  answer("$<2.2 x 10^{-6}$", correct=TRUE),
  answer("0.723"),
  answer("0.728")
)
```

```{r, penguins-10}
question(
  "Is this value significant?",
  answer("YES!", message="Actually, it can't be! We hadn't defined a significance level. A $p$-value can only be said to be significant when compared to a predecided level at which we are prepared to accept mistakes (e.g 0.05 for 1 in 20 miscalls). Fixing this level  _after_ seeing the $p$-value isn't right "),
  answer("No.", message="Right, but not because of the value. It's because we hadn't defined a significance level. A $p$-value can only be said to be significant when compared to a predecided level at which we are prepared to accept mistakes (e.g 0.05 for 1 in 20 miscalls). Fixing this level  _after_ seeing the $p$-value isn't right"),
  answer("Yes, but no.", correct=TRUE, message="This $p$-value would be significant at $0.05$, but we didn't specifiy the significance level before hand, so we don't get it quite right. A $p$-value can only be said to be significant when compared to a predecided level at which we are prepared to accept mistakes (e.g 0.05 for 1 in 20 miscalls). Fixing this level  _after_ seeing the $p$-value isn't right" )
)
```